---
title: "Initial Korean covid 19"
author: "Melanie Vera"
date: "October 10, 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

#Methods

Result - categorical
sex - categorical

We use the cat~cat helper file.

$H_O$: Dying is not independent of sex

$H_A$: Dying is dependent of sex.

## Graphical results - use ggplot commands to do this  


```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```
  
The bar chart shows that men die at a higher rate compared to women.




  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

Among the people who die there was 64% of men who died and 36% of women who died.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The  p value from the chi-square test was 0.001, so we reject the null hypothesis. We conclude that dying is dependent of sex. 2.3% of men are more likly to die of covid 19 in Korea than women in Korea.

## Conclusions
In conclusion, deaths from covid 19 in korea are dependent on sex. In the fisher's exact test, it shows that men are 2.3 times more likely to die from covid than women. We got this from the cat~cat file. The p value was 0.01, meaning that we reject the null hypothesis because it is less than 0.05. 
